---
title: "appendix"
output: html_document
date: '2022-05-07'
---


```{r}
summary(model91)
```

1s(ppt,year)                    45.090     69  294.178  < 2e-16 ***
2s(vpma,year)                   39.153     69  189.674  < 2e-16 ***
3s(perc_cloud)                   7.555      9   53.720  < 2e-16 ***
4s(wind_speed)                   3.411      9   11.064 0.008069 ** 
5s(pFederal)                     5.625      9   77.413  < 2e-16 ***
6s(pState)                       1.649      9   19.540 4.82e-06 ***
7s(dTrail)                       5.255      9   25.204 4.60e-05 ***
8s(dTemp,year)                  53.213     69  134.457  < 2e-16 ***
9s(dPpt,year)                   46.023     64  329.540  < 2e-16 ***
10s(dVPM,year)                   49.558     68  122.717  < 2e-16 ***
11s(dPowerLine)                   1.894      9   23.161 3.09e-06 ***
12s(herb)                         5.719      9  333.307  < 2e-16 ***
13s(elevation)                    6.697      9   31.602 7.21e-06 ***
14s(slope)                        7.273      9   36.948  < 2e-16 ***
15s(CENTROID_X,CENTROID_Y,year) 208.266    209 1801.085  < 2e-16 ***
16s(temp,year)                   52.366     69  221.632  < 2e-16 ***
17s(pPBA,year)                    6.855     66   35.507  < 2e-16 ***
18s(pb1Prev,year)                 3.275     65    4.317 0.202193    
19s(pb2Prev,year)                 6.607     64   13.513 0.038353 *  
20s(pb3Prev,year)                 6.749     65   21.385 0.000431 ***
```{r}
variables = c("precipitation (mm)", "max vpd", "cloud corerage", "wind speed (m/s)", "percentage federal ownership",
              "percentage state ownership", "distance to hiking trail (km)", 
              "deviation from average temperature (degree Celcius)", "deviation from average precipitation (mm)", 
              "deviation from average max vpd", "distance to energy transmission line (km)", 
              "percentage herb coverage", "elevation (m)", "slope (degree)","no", "temperature (degree Celcius)", 
              "percentage prescribed burned", "percentage prescribed burned (1 year prior)",
              "percentage prescribed burned (2 years prior)", "percentage prescribed burned (3 years prior)"
              )
orders = c(1, 9, 16, 8, 2,
           10, 3, 4, 12, 13,
           14, 5, 6, 7, 11,
           17:20)




plots = list()
counter = 0
for (i in orders[1:5]) {
  counter = counter + 1
  p = plot(sm(viz_binary, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}


pg1 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], nrow=1,
                labels = paste("(", letters[1:5], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0)

plots = list()
counter = 0
for (i in orders[6:10]) {
  counter = counter + 1
  p = plot(sm(viz_binary, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}

pg2 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], nrow=1,
                labels = paste("(", letters[6:10], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0)

plots = list()
counter = 0
for (i in orders[11:15]) {
  counter = counter + 1
  p = plot(sm(viz_binary, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}

pg3 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], nrow=1,
                labels = paste("(", letters[11:15], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0)

plots = list()
counter = 0
for (i in orders[16:18]) {
  print(i)
  print(variables[i])
  counter = counter + 1
  p = plot(sm(viz_binary, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}

p1 = plot(sm(viz_binary, select=20)) + 
  l_fitLine() +
  xlab(variables[20]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") 

p1 = p1$ggObj
plots[[4]] = p1

pg4 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], NULL, nrow=1,
                labels = paste("(", letters[16:19], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0, rel_widths = c(1, 1, 1, 1.5, 1))

plot_grid(pg1, pg2, pg3, pg4, nrow=4, align="hv")
ggsave("~/R/ss/Plots/appendix1.png", width = 20, height = 15, units = "in")


```

```{r}
summary(model_cont)
```

1s(ppt,year)                   10.0625     69 18.588 0.002785 ** 
2s(vpma,year)                   2.1917     69  2.195 0.264188    
3s(perc_cloud)                  0.1943      9  0.083 0.481256    
4s(pop_den)                     0.2197      9  0.098 0.566453    
5s(pFederal)                    0.9955      9  6.173 0.004712 ** 
6s(pState)                      0.1652      9  0.005 0.981389    
7s(dTemp,year)                  0.7697     69  0.300 0.975712    
8s(dPpt,year)                   4.6333     69 11.058 0.000457 ***
9s(dVPM,year)                   1.5104     69  1.230 0.403647    
10s(herb)                        1.6921      9  9.483 0.002097 ** 
11s(elevation)                   0.9453      9  1.948 0.082083 .  
12s(slope)                       1.4389      9 10.180 0.000905 ***
13s(CENTROID_X,CENTROID_Y,year)  3.2651    209  2.802 0.447507    
14s(temp,year)                   1.5792     69  1.014 0.695585    
15s(pPBA,year)                   0.5074     66  0.154 0.984202    
16s(pb1Prev,year)                0.4477     67  0.091 0.997632    
17s(pb2Prev,year)                0.4691     67  0.081 0.999477    
18s(pb3Prev,year)                0.4479     67  0.083 0.999362   



```{r}
viz_cont = getViz(model_cont)
```

              )
```{r}
variables = c("precipitation (mm)", "max vpd", "cloud corerage", "population density", "percentage federal ownership",
              "percentage state ownership", "deviation from average temperature (degree Celcius)", 
              "deviation from average precipitation (mm)", "deviation from average max vpd",
              "percentage herb coverage", "elevation (m)", "slope (degree)","no",
              "temperature (degree Celcius)", 
              "percentage prescribed burned", "percentage prescribed burned (1 year prior)",
              "percentage prescribed burned (2 years prior)", "percentage prescribed burned (3 years prior)")

orders = c(1, 8, 14, 7, 2,
          9, 3, 10, 11, 12,
          4, 5, 6, 15, 16,
          17, 18)

plots = list()
counter = 0
for (i in orders[1:5]) {
  counter = counter + 1
  p = plot(sm(viz_cont, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}


pg1 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], nrow=1,
                labels = paste("(", letters[1:5], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0)

plots = list()
counter = 0
for (i in orders[6:10]) {
  print(i)
  counter = counter + 1
  p = plot(sm(viz_cont, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}

pg2 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], nrow=1,
                labels = paste("(", letters[6:10], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0)

plots = list()
counter = 0
for (i in orders[11:15]) {
  counter = counter + 1
  p = plot(sm(viz_cont, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}

pg3 = plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], nrow=1,
                labels = paste("(", letters[11:15], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0)

plots = list()
counter = 0
for (i in orders[16]) {
  counter = counter + 1
  p = plot(sm(viz_cont, select=i)) + 
  l_fitLine() +
  xlab(variables[i]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") +
  theme(legend.position="none")
  
  
  plots[[counter]] = p$ggObj
}

p1 = plot(sm(viz_cont, select=18)) + 
  l_fitLine() +
  xlab(variables[18]) +
  ylab("log-odds") +
  labs(color="year") +
  #ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal(base_size = 10, base_family="Times New Roman") 

p1 = p1$ggObj
plots[[2]] = p1

pg4 = plot_grid(plots[[1]], plots[[2]],NULL, NULL, NULL, nrow=1,
                labels = paste("(", letters[16:17], ")", sep=""), label_size=10, label_fontfamily = "Times New Roman", hjust=0, rel_widths = c(1, 1.5, 1, 1, 1))

plot_grid(pg1, pg2, pg3, pg4, nrow=4, align="hv")
ggsave("~/R/ss/Plots/appendix2.png", width = 20, height = 15, units = "in")
```


