---
title: "mixed_effects_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="/home/michael_xiao/T/st")
getwd()
```






```{r}
# load packages and data
require(sf)
require(sp)
require(tidyverse)
require(ggplot2)
require(plotly)
require(mgcv)
require(GGally)
require(car)
require(gstat)
require(iml)
require(ggthemes)
require(mgcViz)
require(tidymv)
require(gamm4)
require(ggforce)
require(gratia)


data = read_sf("/home/michael_xiao/R/ss/Final_Data/update_data2.shp")
data = data %>% filter((tree > 0.5)|(shrub > 0.6))

data = dplyr::rename(data, y2010.pPBA=pPBA10, y2011.pPBA=pPBA11, y2012.pPBA=pPBA12, y2013.pPBA=pPBA13, 
                       y2014.pPBA=pPBA14, y2015.pPBA=pPBA15, y2016.pPBA=pPBA16, y2017.pPBA=pPBA17, 
                       y2018.pPBA=pPBA18, y2019.pPBA=pPBA19,
                       y2010.wf = wf2010, y2011.wf = wf2011, y2012.wf = wf2012, y2013.wf = wf2013, 
                       y2014.wf = wf2014, y2015.wf = wf2015, y2016.wf = wf2016, y2017.wf = wf2017, 
                       y2018.wf = wf2018, y2019.wf = wf2019, 
                       y2011.pb = pb2011, y2012.pb = pb2012, y2013.pb = pb2013, y2014.pb = pb2014, 
                       y2015.pb = pb2015, y2016.pb = pb2016, y2017.pb = pb2017, y2018.pb = pb2018, 
                       y2019.pb = pb2019, y2010.pb = pb2010, 
                       y2010.pWFA = pWFA2010, y2011.pWFA = pWFA2011, y2012.pWFA = pWFA2012, y2013.pWFA = pWFA2013, 
                       y2014.pWFA = pWFA2014, y2015.pWFA = pWFA2015, y2016.pWFA = pWFA2016, y2017.pWFA = pWFA2017, 
                       y2018.pWFA = pWFA2018, y2019.pWFA = pWFA2019,
                       y2010.temp = Avg_y2010_, y2011.temp = Avg_y2011_, y2012.temp = Avg_y2012_, y2013.temp = Avg_y2013_, 
                       y2014.temp = Avg_y2014_, y2015.temp = Avg_y2015_, y2016.temp = Avg_y2016_, y2017.temp = Avg_y2017_, 
                       y2018.temp = Avg_y2018_, y2019.temp = Avg_y2019_, 
                       precip = ppt, temperature = temp)
data
```

```{r}
# read in fire data
fire = read_sf("/home/michael_xiao/R/s/fire_data/fire20_1.gdb", layer="firep20_1")
fire$YEAR_ = as.numeric(fire$YEAR_)
fire = fire %>% filter((YEAR_ >= 2010) & (YEAR_ <= 2019))

# clip fire

```

```{r}
# encode lags
for (y in 2013:2019) {
  data[,paste("y", y, ".pb1Prev", sep="")] = data[,paste("y", y-1, ".pPBA", sep="")] %>% st_drop_geometry() %>% pull()
  data[,paste("y", y, ".pb2Prev", sep="")] = data[,paste("y", y-2, ".pPBA", sep="")] %>% st_drop_geometry() %>% pull()
  data[,paste("y", y, ".pb3Prev", sep="")] = data[,paste("y", y-3, ".pPBA", sep="")] %>% st_drop_geometry() %>% pull()
}

# shift aspect
data$aspect = abs(180 - abs(data$aspect-225))

# 
data.cent = data$geometry %>% st_centroid()
new.proj = st_transform(data.cent, crs=4326) %>% st_coordinates()
data$CENTROID_X = new.proj[,1]
data$CENTROID_Y = new.proj[,2]
data
```




```{r}
# drop monthly wf, pb, wf_count, pb_count, annual_wf, sd_wf, nMonth_wf, annual_pb, sd_pb, wf
drops = c(paste("wf", 1:12, sep=""), paste("pb", 1:12, sep=""), "wf_count", "pb_count", "annual_wf", 
          "sd_wf", "nMonth_wf", "annual_pb", "sd_pb","nMonth_pb", "pb_score", "fire_score")
data = data[,!names(data) %in% drops]
data = st_drop_geometry(data)

# reshape data
dataLong = data %>% 
  pivot_longer(starts_with("y"), names_to=c("year", ".value"), names_pattern="y(201\\d)[\\.|_](.*)")

dataLong = dataLong %>% filter(year >= 2013)
dataLong$year = as.factor(dataLong$year)
dataLong$wf = as.numeric(dataLong$pWFA > 0)

# encode deviances
dataLong$dTemp = dataLong$temp - dataLong$temperature
dataLong$dPpt = dataLong$ppt - dataLong$precip
dataLong$dVPM = dataLong$vpma - dataLong$vpmax

dataLong
```





"s(ppt) + s(temp) + s(vpmax) + s(vpmin) + s(elevation) + s(perc_cloud) + s(wind_speed) + s(pop_den) +
    s(pFederal) + s(pState) + s(pCounty) + s(pCity) + s(pNP) + s(pSD) + s(pPrivate) + s(dRoad) + s(dTrail) +
    s(dPowerLine) + s(shrub) + s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X) + s(CENTROID_Y) + year + s(pPBA)"
    
  "ppt + temp + vpmax + vpmin + elevation + perc_cloud + wind_speed + pop_den + pFederal + pState + pCounty + pCity + pNP + pSD + pPrivate + dRoad + dTrail + dPowerLine + shrub + tree + herb + slope + aspect + CENTROID_X + CENTROID_Y + year + pPBA"

```{r}
# vif analysis
# removed: vpmax, all but pFederal and pState, elevation, shrub
vif_form = formula(paste("wf~", "ppt + temp + vpmin  + perc_cloud + wind_speed + pop_den + pFederal + pState +dRoad + dTrail + dPowerLine + tree + herb + slope + aspect + CENTROID_X + CENTROID_Y + pPBA", sep=""))
vif_model = lm(vif_form, data=dataLong)
car::vif(vif_model)
```





```{r}
# fit naive gam: make formula
naive_vars = colnames(dataLong)[-which(colnames(dataLong) == "wf" | colnames(dataLong) == "pWFA" | colnames(dataLong) == "pb")]
for_pasting = paste(paste("s(", naive_vars, ")", sep=""), collapse=" + ")
naive_form = formula(paste("wf~",   "s(ppt, bs='re') + s(vpma, bs='re') + s(perc_cloud) + s(wind_speed) + 
s(pop_den) +
    s(pFederal) + s(pState) + s(dRoad) + s(dTrail) +
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + 
    s(year, bs='re') + s(year, temp, bs='re') + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev)", sep=""))

naive_form
```
```{r}
# fit naive gam: fitting
naive_gam = gam(naive_form, data=dataLong, family=binomial)
summary(naive_gam)
```

```{r}
naive_gam$coefficients
```


```{r}
viz = getViz(naive_gam)
plot(sm(viz, select=1))
```






```{r}
plot(naive_gam, scale=0, rug=T)
```

Model 1: only random intercept on year (G Model)
```{r}
model1_form = formula(paste("wf~",   "s(ppt) + s(vpma) + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) +
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + 
    s(year, bs='re') + s(temp) + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev)", sep=""))

model1 = gam(model1_form, data=dataLong, family=binomial)
summary(model1)
```


Model 2: add anomoly detection (G Model)
```{r}

model2_form = formula(paste("wf~",   "s(ppt) + s(vpma) + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + 
    s(year, bs='re') + s(temp) + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev)", sep=""))

model2 = gam(model2_form, data=dataLong, family=binomial)
summary(model2)
```


Model 3: Random slope (Linear?)
```{r}
model3_form = formula(paste("wf~",   "s(ppt, year, bs='re') + s(vpma, year, bs='re') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + s(temp, year, bs='re') + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev) + s(ppt) + s(vpma) + s(temp)", sep=""))

model3 = gam(model3_form, data=dataLong, family=binomial)
summary(model3)
```


Model 4: Random Intercept + Slope
```{r}
model4_form = formula(paste("wf~",   "s(ppt, year, bs='re') + s(vpma, year, bs='re') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + s(temp, year, bs='re') + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev) + s(ppt) + s(vpma) + s(temp) + s(year, bs='re')", sep=""))

model4 = gam(model4_form, data=dataLong, family=binomial)
summary(model4)
```


Model 5: Random Smooth (GS Model)
```{r}
model5_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + s(temp, year, bs='fs') + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev) + s(ppt) + s(vpma) + s(temp)", sep=""))

model5 = gam(model5_form, data=dataLong, family=binomial)
summary(model5)
```

Model 6: Additional Smoothers on Fire Stuff (GS Model)
```{r}
model6_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs') + s(pb2Prev, year, bs='fs') + s(pb3Prev, year, bs='fs') + s(ppt) + s(vpma) + s(temp)", sep=""))

model6 = gam(model6_form, data=dataLong, family=binomial)
summary(model6)
```

Model 7: GI Model (different penalty terms)

```{r}
model7_form = formula(paste("wf~",   "s(ppt, by=year) + s(vpma, by=year) + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + s(temp, by=year) + s(pPBA, by=year) + s(pb1Prev, by=year) + s(pb2Prev, by=year) + s(pb3Prev, by=year) + s(ppt) + s(vpma) + s(temp) + s(year, bs='re')", sep=""))

model7 = gam(model7_form, data=dataLong, family=binomial, method="REML")
summary(model7)
```

Model 8: Cyclic for aspect (GS Model)

```{r}
model8_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp) + s(dPpt) + s(dVPM)+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect, bs='cc') + s(CENTROID_X, CENTROID_Y) + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs') + s(pb2Prev, year, bs='fs') + s(pb3Prev, year, bs='fs') + s(ppt) + s(vpma) + s(temp)", sep=""))

ctrl <- gam.control(trace = TRUE)
model8 = gam(model8_form, data=dataLong, family=binomial, knots=list(aspect=c(0,360)), control = ctrl)
summary(model8)
```

Model 9: GS Model with deviance factored
```{r}
model9_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp, year, bs='fs') + s(dPpt, year, bs='fs') + s(dVPM, year, bs='fs') +
    s(dPowerLine) +  s(tree) + s(herb)+ s(elevation) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y, year, bs='fs') + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs') + s(pb2Prev, year, bs='fs') + s(pb3Prev, year, bs='fs') + s(ppt) + s(vpma) + s(temp) + s(dTemp) + s(dPpt) + s(dVPM)", sep=""))

model9 = gam(model9_form, data=dataLong, family=binomial)


summary(model9)


```

Model 10: Tensor product of year and coords

```{r}
model10_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp, year, bs='fs') + s(dPpt, year, bs='fs') + s(dVPM, year, bs='fs')+
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + te(CENTROID_X, CENTROID_Y, year, bs='fs') + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs') + s(pb2Prev, year, bs='fs') + s(pb3Prev, year, bs='fs') + s(ppt) + s(vpma) + s(temp) + s(dTemp) + s(dPpt) + s(dVPM)", sep=""))

model10 = bam(model10_form, data=dataLong, family=binomial())

saveRDS(model10, file = "/home/michael_xiao/R/ss/models/model10.rds")
```

Model 11: S model
```{r}
model11_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp, year, bs='fs') + s(dPpt, year, bs='fs') + s(dVPM, year, bs='fs') +
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y, year, bs='fs') + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs') + s(pb2Prev, year, bs='fs') + s(pb3Prev, year, bs='fs')", sep=""))

model11 = bam(model11_form, data=dataLong, family=binomial)
summary(model11)
saveRDS(model11, file = "/home/michael_xiao/R/ss/models/model11.rds")
```



```{r}
concurvity(model9)
```

```{r}
dataLong = as.data.frame(dataLong)
ggplot(data=dataLong) +
  geom_line(aes(x=index2, y=residual), alpha=0.5)
```


```{r}
plot(model11,scale=0)
```



```{r}
res = model9$residuals
dataLong$res = res
x = dataLong$CENTROID_X
y = dataLong$CENTROID_Y
coordinates(dataLong) = ~x+y
v = variogram(res~1, data=dataLong)

v_dir = variogram(res~1, data=dataLong, alpha=c(0,45,90,130))
plot(v)
plot(v_dir)
plot_map(data %>% filter(res>0), "res", direction=-1)
```


```{r}
plot_smooths(model9, series=pPBA, exclude_random=T, comparison=year)
```




Model Comparison

```{r}
AIC(model1, model2, model3, model4, model5, model6, model7, model9, model10, model91)
```

```{r}
viz = getViz(model9)
plot(sm(viz, select=20), rug=T) + 
  l_fitLine()

#plot(model9, select=11, scale=0, rug=T)
```


```{r}
gam.check(model9)
```



```{r}
plot(model9,scale=0, rug=T)
```


```{r}
dataLong$residual = residuals.gam(model9)
boxplot(residual~year, data=dataLong, outline=F)
```














```{r}
# mapping the variables
plot_map = function(data, var, type="c", direction=1) {
  if (type == "c") {
    ggplot(data=data) +
      geom_sf(aes(fill=!!sym(var), color=!!sym(var))) +
      scale_fill_viridis_c(direction=direction) +
      scale_color_viridis_c(direction=direction)
  } else if (type == "d") {
    ggplot(data=data) +
      geom_sf(aes(fill=as.factor(!!sym(var)), color=as.factor(!!sym(var)))) +
      scale_fill_viridis_d(direction=direction) +
      scale_color_viridis_d(direction=direction)
  }
}

plotFireVar = function(data, var, plot_title, legend_title) {
  ggplot() +
  geom_sf(data=data, aes(color=!!sym(var))) +
    scale_color_continuous(high = "#132B43", low = "#56B1F7", name=legend_title) +
  geom_sf(data=fire_loc, size=0.4, alpha=0.15, aes(fill="val"), color="red") +
    scale_fill_manual(values="black", labels="", guide=guide_legend(override.aes=list(size=2, alpha=1), title="Fire")) +
  ggtitle(plot_title)+
  theme_minimal()
}



dataPlot = read_sf("/home/michael_xiao/R/ss/Final_Data/update_data2.shp")
dataPlot = dataPlot %>% filter((tree > 0.5)|(shrub > 0.6))

# make fire location
fire_loc = dataPlot %>% filter(wf_count >= 1) %>% pull(geometry) %>% st_centroid()
fire_loc = st_sf(fire_loc)
fire_loc$val = 1.5

# test on dtrail
ggplot() +
  geom_sf(data=dataPlot, aes(color=dTrail)) +
    scale_color_continuous(high = "#132B43", low = "#56B1F7", name="Distance(km)") +
  geom_sf(data=fire_loc, size=0.4, alpha=0.15, aes(fill="val"), color="red") +
    scale_fill_manual(values="black", labels="", guide=guide_legend(override.aes=list(size=2, alpha=1), title="Fire")) +
  ggtitle("Distance to Trails")+
  theme_minimal()


plotFireVar(dataPlot, "dTrail", plot_title = "Distance to Trail", legend_title = "Distance(km)")
ggsave("~/R/s/Plots/fire_var/dTrail.png")

colnames(dataPlot)

vars_to_plot = c("ppt", "temp", "vpmax", "elevation", "perc_cloud", "wind_speed", "pop_den", "pFederal", "pState", "dRoad", "dTrail", "dPowerLine", "annual_pb", "slope", "aspect")

plot_titles = c("Mean Annual Precipitation", "Mean Annual Temperature", "Mean Max VPD", "Elevation", "Cloud Coverage", "Wind Speed", "Population Density", "Proportion of Federally Owned Land", "Proportion of State Owned Land", "Distance to Road", "Distance to Trail", "Distance to Power Line", "Mean Annual Proportion Prescribed Burning", "Slope", "Aspect")

legend_titles = c("Precipitation(mm)", "Temperature(degree Celcius)", "VPD(hPA)", "Elevation(m)", "Cloud Coverage(proportion)", "Wind Speed(m/s)", "Population Density(persons/cell)", "Proportion of Federally Owned Land", "Proportion of State Owned Land", "Distance(km)", "Distance(km)", "Distance(km)", "Mean Annual Proportion Prescribed Burning", "Slope(degree)", "Aspect(folded degree)")

for (var in vars_to_plot) {
  print(var)
  pt = plot_titles[which(vars_to_plot == var)]
  lt = legend_titles[which(vars_to_plot == var)]
  plotFireVar(dataPlot, var, plot_title = pt, legend_title = lt)
  ggsave(paste("~/R/s/Plots/fire_var/", var, ".png"))
}
```


```{r}
gam.check(model9)
```

```{r}
# increase k
model7_form = formula(paste("wf~",   "s(ppt, by=year) + s(vpma, by=year) + s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail) + s(dTemp, by=year) + s(dPpt, by=year) + s(dVPM, by=year)+
    s(dPowerLine) +  s(tree) + s(herb) + s(elevation) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y, by=year) + s(temp, by=year) + s(pPBA, by=year) + s(pb1Prev, by=year) + s(pb2Prev, by=year) + s(pb3Prev, by=year) + s(ppt) + s(vpma) + s(temp) + s(dTemp) + s(dPpt) + s(dVPM) + s(year, bs='re')", sep=""))

model7 = bam(model7_form, data=dataLong, family=binomial())
saveRDS(model7, file = "/home/michael_xiao/R/ss/models/model7.rds")

summary(model7)

plot(model7,scale=0)
```



```{r}
plot(model92, scale=0, rug=T)
```

```{r}
summary(model9)
```

```{r}
# continuous model
# wind speed, dTrail, dPowerLine, tree, 
form_cont = formula(paste("pWFA~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(pop_den) + s(pFederal) + s(pState) + s(dTemp, year, bs='fs') + s(dPpt, year, bs='fs') + s(dVPM, year, bs='fs') + s(herb) + s(elevation) + s(slope) + s(elevation) + s(CENTROID_X, CENTROID_Y, year, bs='fs') + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs')  + s(pb2Prev, year, bs='fs')  + s(pb3Prev, year, bs='fs')", sep=""))

model_cont = gam(form_cont, data=dataLong, select=T, family=betar())
saveRDS(model_cont, file = "/home/michael_xiao/R/ss/models/model_cont.rds")

summary(model_cont)
```

```{r}
cont = plot(model_cont, scale=0, select=17)

percent_diff = diff(cont[[6]]$fit) *100 / cont[[6]]$fit[1:(length(cont[[6]]$fit)-1)]

mean(percent_diff/ diff(cont[[6]]$x))

plot(model_cont, scale=0)

plot_smooth(model91, view="pPBA", plot_all = "year", rm.ranef = F, shade=F, se=0)
```


```{r}
dataLong
```

```{r}
data9 = plot(model9, select=20, se=T, scale=0)
```

```{r}
data9
```

```{r}
k.check(model9)
```

```{r}
gamm_form = formula(paste("wf~",   "s(perc_cloud) + s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dRoad) + s(dTrail)  +
    s(dPowerLine) +  s(tree) + s(herb) + s(slope) + s(aspect) + s(CENTROID_X, CENTROID_Y) + s(pPBA) + s(pb1Prev) + s(pb2Prev) + s(pb3Prev, year) + s(ppt) + s(vpma) + s(temp) + s(dTemp) + s(dPpt) + s(dVPM)", sep=""))
#dataLong$year = as.numeric(dataLong$year) +2012

model_gamm = gamm4(gamm_form, data=dataLong, family=binomial(), random=~(1|year))
saveRDS(model_gamm, file = "/home/michael_xiao/R/ss/models/model_gamm.rds")

dataLong
```

```{r}
dataPlot = read_sf("/home/michael_xiao/R/ss/Final_Data/update_data2.shp")
dataPlot = dataPlot %>% filter((tree > 0.5)|(shrub > 0.6))
ggplot() +
  geom_sf(data=dataPlot, fill="#ead47f", color="black", lwd=0.05) +
  ggtitle("Grids for Analysis")+
  theme_minimal()

nrow(dataPlot)
```







```{r}
ggplot() +
  geom_sf(data=dataPlot, aes(color=dTrail)) +
    scale_color_continuous(high = "#132B43", low = "#56B1F7", name="Distance(km)") +
  geom_sf(data=fire_loc, size=0.4, alpha=0.15, aes(fill="val"), color="red") +
    scale_fill_manual(values="black", labels="", guide=guide_legend(override.aes=list(size=2, alpha=1), title="Fire")) +
  ggtitle("Distance to Trails")+
  theme_minimal()

plot_map = function(data, var, type="c", direction=1, title="", legend_title) {
  if (type == "c") {
    ggplot(data=data) +
      geom_sf(aes(fill=!!sym(var), color=!!sym(var))) +
      scale_fill_viridis_c(direction=direction, name=legend_title) +
      scale_color_viridis_c(direction=direction, guide="none") +
      # scale_color_gradient(low="grey", high="red") +
      # scale_fill_gradient(low="grey", high="red") +
      ggtitle(title) +
      theme_minimal()
  } else if (type == "d") {
    ggplot(data=data) +
      geom_sf(aes(fill=as.factor(!!sym(var)), color=as.factor(!!sym(var)))) +
      scale_fill_manual(labels=c("No", "Yes"), values=c("gray", "red"), name=legend_title) +
      scale_color_manual(labels=c("No", "Yes"), values=c("gray", "red"), guide="none") + 
      ggtitle(title) +
      theme_minimal()
  }
}
```


```{r}
plot_map(dataPlot, "pPBA19",direction=-1, type="c", title="% Prescribe Burned (2019)", legend_title="percentage")

ggplot(data=dataPlot, aes(x=pPBA19^(1/3), fill=factor(pPBA19^(1/3)))) +
  geom_histogram(show.legend = F, binwidth = 0.001) +
  scale_fill_viridis_d() + 
  ggtitle("Distribution of Temperature") +
  xlab("temperature") +
  theme_minimal()

hist(dataPlot$temp)

colnames(dataPlot)


ggplot(data=dataPlot, aes(x=pPBA13, fill=factor(pPBA13))) +
  geom_histogram(show.legend = F) +
  facet_zoom(x=pPBA13>0, ylim=c(0,20)) +
  scale_fill_viridis_d(direction=-1) + 
  ggtitle("Distribution of % PB (2013)") +
  xlab("percentage")

dataPlot

ggplot(data=dataPlot, aes(x=Avg_y2018_)) +
  geom_histogram() 

ppba = dataPlot[paste("pPBA", 13:19, sep="")] %>% st_drop_geometry()
ppba_count = apply(ppba, MARGIN=2, FUN=function(x){return (sum(x>0.2))})

ppba_count = data_frame(year=2013:2019, count=ppba_count)
ppba_count

ggplot(data=ppba_count) +
  geom_col(aes(x=factor(year), y=count), fill="blue", alpha=0.7) +
  ggtitle("Number of Large Scale (>25%) Wildfires by Year") + 
  scale_y_continuous(n.breaks=7) +
  xlab("year") +
  theme_minimal()

dataPlot
```

```{r}
ggplot(data=dataPlot, aes(x=Avg_y2017_, fill=factor(Avg_y2017_))) +
  geom_histogram(show.legend = F) +
  facet_zoom(x=pPBA13>0, ylim=c(0,20)) +
  scale_fill_viridis_d(direction=-1) + 
  ggtitle("Distribution of % PB (2013)") +
  xlab("percentage")

dataPlot
```





```{r}
model91_form = formula(paste("wf~",   "s(ppt, year, bs='fs') + s(vpma, year, bs='fs') + s(perc_cloud) + s(wind_speed) + s(pFederal) + s(pState) + s(dTrail) + s(dTemp, year, bs='fs') + s(dPpt, year, bs='fs') + s(dVPM, year, bs='fs') +
    s(dPowerLine) +  s(herb)+ s(elevation) + s(slope) + s(CENTROID_X, CENTROID_Y, year, bs='fs') + s(temp, year, bs='fs') + s(pPBA, year, bs='fs') + s(pb1Prev, year, bs='fs') + s(pb2Prev, year, bs='fs') + s(pb3Prev, year, bs='fs')", sep=""))

model91 = gam(model91_form, data=dataLong, family=binomial, select=T)
saveRDS(model91, file = "/home/michael_xiao/R/ss/models/model91.rds")

summary(model91)
```

```{r}
toy_x = runif(n=20, min=-10, max=10)
toy_y = toy_x^3 + rnorm(n=20, mean=0, sd=10)
df = data.frame(x=toy_x, y=toy_y)

ggplot(data=df, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method="lm", formula=y~1+x, se=F)

ggplot(data=df, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method="gam", formula=y~1+s(x), se=F)

```

wf ~ s(ppt, year, bs = "fs") + s(vpma, year, bs = "fs") + s(perc_cloud) + 
    s(wind_speed) + s(pFederal) + s(pState) + s(dTrail) + s(dTemp, 
    year, bs = "fs") + s(dPpt, year, bs = "fs") + s(dVPM, year, 
    bs = "fs") + s(dPowerLine) + s(herb) + s(elevation) + s(slope) + 
    s(CENTROID_X, CENTROID_Y, year, bs = "fs") + s(temp, year, 
    bs = "fs") + s(pPBA, year, bs = "fs") + s(pb1Prev, year, 
    bs = "fs") + s(pb2Prev, year, bs = "fs") + s(pb3Prev, year, 
    bs = "fs")



pWFA ~ s(ppt, year, bs = "fs") + s(vpma, year, bs = "fs") + s(perc_cloud) + 
    s(wind_speed) + s(pop_den) + s(pFederal) + s(pState) + s(dTrail) + 
    s(dTemp, year, bs = "fs") + s(dPpt, year, bs = "fs") + s(dVPM, 
    year, bs = "fs") + s(dPowerLine) + s(tree) + s(herb) + s(elevation) + 
    s(slope) + s(elevation) + s(CENTROID_X, CENTROID_Y, year, 
    bs = "fs") + s(temp, year, bs = "fs") + s(pPBA, year, bs = "fs") + 
    s(pb1Prev, year, bs = "fs")

```{r}
cont_vars = c("ppt", "vpma", "perc_cloud", "wind_speed", "pop_den", "pFederal", "pState", "dTrail", "dTemp", "dPpt", "dVPM", "dPowerLine","tree", "herb", "elevation", "slope", "CENTROID_X", "CENTROID_Y", "year", "pPBA","temp", "pb1Prev")

X_cont = dataLong[cont_vars]
y_cont = dataLong$pWFA

mod_cont = Predictor$new(model_cont, data=X_cont, y=y_cont)
imp_cont = FeatureImp$new(mod_cont, loss = "mse")

saveRDS(imp_cont, file="~/R/ss/models/imp_cont.rds")


labels_cont = c("y coordiante", "x coordinate", "year", "max vpd deviation", "temperature", "max vpd", "% federal owned", "ppt deviation", "ppt", "% herb", "% tree", "slope", "elevation", "dist powerline", "% cloud", "temperature deviation", "% pb", "% pb 1 year prior", "dist trail", "population density", "wind speed", "% state owned")

plot(imp_cont) + 
  scale_y_discrete(label=rev(labels_cont))


```





```{r fig.height=5}
x_vars = c("temp", "ppt", "vpma", "perc_cloud", "wind_speed", "pFederal", "pState", "dTrail", "dTemp", "dPpt", "dVPM", "dPowerLine", "herb", "elevation", "slope", "CENTROID_X", "CENTROID_Y", "year", "pPBA", "pb1Prev", "pb2Prev", "pb3Prev")
X = dataLong[x_vars]
y = dataLong$wf

mod <- Predictor$new(model91, data = X, y = y)

imp <- FeatureImp$new(mod, loss = "mse")



saveRDS(imp, file="~/R/ss/models/imp_bi.rds")

labels_bin = c("y coordiante", "x coordinate", "% federal owned", "temperature", "% cloud", "elevation", "temperature deviation", "dist trail", "% state owned", "% pb burned", "% pb 3 years ago", "% pb 2 years ago", "% pb 1 year ago", "wind speed", "slope", "ppt deviation", "max vpd deviation", "year", "% herb", "ppt", "max vpd", "dist powerline")

plot(imp) + 
  facet_zoom(xlim=c(0.9925, 1.005)) + 
  scale_y_discrete(label=rev(labels_bin)) +
  theme(text = element_text(size=15))

ggsave(filename="~/R/ss/Plots/imp_bin.png", width=7, height=10, units="in")
```

```{r}
labels_bin = c("y coordiante", "x coordinate", "% federal owned", "% cloud", "elevation", "temperature deviation", "dist trail", "% state owned", "% pb burned", "% pb 3 years ago", "% pb 2 years ago", "% pb 1 year ago", "wind speed", "slope", "ppt deviation", "max vpd deviation", "year", "% herb", "ppt", "max vpd", "dist powerline")

```

Approximate significance of smooth terms:
                                    edf Ref.df   Chi.sq  p-value    
s(ppt,year)                    21.47979     29  112.333  < 2e-16 ***
s(vpma,year)                   23.67795     28  167.723  < 2e-16 ***
s(perc_cloud)                   7.80658      9   40.793  < 2e-16 ***
s(wind_speed)                   2.30855      9    9.539 0.005310 ** 
s(pFederal)                     6.15548      9  145.901  < 2e-16 ***
s(pState)                       3.61030      9   25.985 8.70e-06 ***
s(dTrail)                       6.35781      9   30.122 1.29e-05 ***
s(dTemp,year)                  23.91842     28   64.701 4.46e-06 ***
s(dPpt,year)                   18.84981     28  125.773  < 2e-16 ***
s(dVPM,year)                   18.94525     28  127.812  < 2e-16 ***
s(dPowerLine)                   7.79829      9   28.504 0.000275 ***
s(herb)                         5.98996      9  368.864  < 2e-16 ***
s(elevation)                    6.57144      9   29.943 1.80e-05 ***
s(slope)                        6.91942      9   36.108  < 2e-16 ***
s(CENTROID_X,CENTROID_Y,year) 107.43797    108 1623.482  < 2e-16 ***
s(temp,year)                   26.39545     28  221.274  < 2e-16 ***
s(pPBA,year)                   11.32956     28   48.894  < 2e-16 ***
s(pb1Prev,year)                 0.00536     28    0.007 0.254511    
s(pb2Prev,year)                 0.84495     28    3.331 0.046703 *  
s(pb3Prev,year)                12.12403     28   23.244 0.002005 ** 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

```{r}
concurvity(model_cont)
```

```{r}
summary(model91)
```


```{r}
viz_binary = getViz(model91)
viz_cont = getViz(model_cont)

p = plot(sm(viz_binary, select=17)) + 
  l_fitLine() +
  #xlab("% pb burned") +
  ylab("log-odds") +
  ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal() 
    
p1 = ggplot() + geom_point(x=x, y=y)

p %+% geom_point(data=dat, aes(x=x, y=y))

print(p1)
  
ggplot() +
  geom_point(aes(x=x,y=y))

ggplot() +
  geom_point(aes(x=dataLong$pPBA[which(dataLong$pPBA >= 0.25)], y=val[which(dataLong$pPBA >= 0.25), "s(pPBA,year)"]))
draw(model91, select=16) 

viz_binary

o = plot(sm(viz_binary, select=19))
listLayers(o)
plot(model_cont, select=1, scale=0)
summary(model91)

plot_smooth(model91, view="pPBA", plot_all = "year", rm.ranef = F, shade=F)

listLayers(o)

bin = plot(model91, se=T, scale=0)
bin[[1]]

val = predict(model91, type = "terms")
y = val[which(dataLong$pPBA >= 0.25)]
x = dataLong$pPBA[which(dataLong$pPBA >= 0.25)]
year = dataLong$year[which(dataLong$pPBA >= 0.25)]

dat = data.frame(x=x, y=y, id=year)

typeof(p)

p$ggObj + geom_point(data=dat, aes(x=x, y=y))

dat %>% filter(id=="2014")

plot(model_cont, scale=0, rug=T)

plot(sm(viz_cont, select=10)) + 
  l_fitLine() +
  #xlab("% pb burned") +
  ylab("log-odds") +
  ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal() 

hist(dataLong[dataLong$year == "2015", "ppt"] %>% pull())
y

 plot(sm(viz_binary, select=)) + 
  l_fitLine() +
  #xlab("% pb burned") +
  ylab("log-odds") +
  ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal() 
```

```{r}
cont_fit = predict(model_cont, type="terms")
x = dataLong$ppt[dataLong$year=="2015"]
y = cont_fit[which(dataLong$year == "2015", "s(ppt,year)")]

dat_cont = data.frame(x=x, y=y, id="2015")

p_cont = plot(sm(viz_cont, select=1)) + 
  l_fitLine() +
  #xlab("% pb burned") +
  ylab("log-odds") +
  ggtitle("Fitted Effect of Prescribed Burning (Chance of Ignition)") +
  theme_minimal() 

p_cont$ggObj %+% geom_point(data=dat_cont, aes(x=x, y=y))

dat_cont
```




```{r}
dataLong[which(dataLong$pPBA == 0.958264),]

hist(dataLong[dataLong["year"] == 2013, "ppt"] %>% pull())
dataLong[dataLong["year"] == 2013, "precip"] %>%
```


```{r}
covariates = colnames(data)[c(13:30, 51:53, 64:65,118:119)]

wf = data[paste("y20", 10:19, ".pWFA", sep="")]
wf_mean = apply(wf, MARGIN=1, FUN=mean)
pb = data[paste("y20", 10:19, ".pPBA", sep="")]
pb_mean = apply(pb, MARGIN=1, FUN=mean)


data_cor = data[covariates]
data_cor$pb = pb_mean

correlation = apply(data_cor, MARGIN=2, FUN=function(x) {return (cor(x=wf_mean, y=x, method="pearson"))})

correlation
```

```{r}
gam.check(model_cont)
```




linear: pFederal, dPowerLine, ppt, 

```{r}
plot5 = plot(model91, scale=0, rug=T)

gam.check(model91)

plot(x=plot5[[5]]$x, y=plot5[[5]]$fit)

percent_diff = diff(plot5[[11]]$fit) *100 / plot5[[5]]$fit[1:(length(plot5[[11]]$fit)-1)]

mean(percent_diff/ diff(plot5[[11]]$x))

plot5[[11]]$x
plot5[[11]]$fit

# larger: ppt, temp
# smaller: wind speed, dTrail, dPowerLine, tree, 

# cloud, wind speed, pop_den, pState, dTrail, 
plot(model_cont, scale=0)
```







